Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-06-07 parallel-tractability-J-v9/parallelTractabilityDatalog.tex
revision

\begin{definition}\label{def:ptd}
\textbf{(Parallel Tractability Class)} Given a class $\mathcal{D}$ of
datalog programs sharing the same rule set, we say that $\mathcal{D}$
is a class of datalog programs with \emph{p}arallel
\emph{t}ractability, short a \emph{\texttt{PTD} class},
if there exists an \texttt{NC} algorithm that performs sound and complete materialization for each datalog program
in $\mathcal{D}$. The corresponding class of ontologies of $\mathcal{D}$ is called a
class of ontologies with parallel tractability, short a \texttt{PTO} class.
\end{definition}

@@

\begin{definition}
\textbf{(Materialization Graph)}\label{def:mg}
A \emph{materialization graph}, with respect to
a datalog program $P=\langle R, \textbf{I}\rangle$, is a directed acyclic graph
denoted by $\mathcal{G}=\langle V, E\rangle$ where,
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item $V$ is the node set and $V\subseteq T_R^{\omega}(\textbf{I})$;
  \item $E$ is the edge set and $E\subseteq T_R^{\omega}(\textbf{I})\times T_R^{\omega}(\textbf{I})$.
\end{enumerate}
For each edge of the form $e(v_1, v_2)$ where $v_1$ and $v_2$ are two nodes,
we say that the node $v_1$ is the parent (node) of $v_2$ and the node $v_2$ is the child (node)
of $v_1$. Further, $\forall H,B_1,\ldots,B_n\in V$, $\mathcal{G}$ satisfies the following conditions:
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item $H$ has a parent node or a child node;
  \item if $H$ has the in-degree of $0$, then $H$ is an original fact of $P$;
  \item $B_1,\ldots,B_n\rightarrow H\in P^*$ is satisfied if $e(B_1, H),\ldots,e(B_n, H)\in E$ and
  $B_1,\ldots,B_n$ are all the parents of $H$.
\end{enumerate}
\end{definition}

@@

We say that a materialization graph $\mathcal{G}$ is a \emph{complete materialization graph}
when $\mathcal{G}$ contains all ground atoms in $T_R^{\omega}(\textbf{I})$.
The set of nodes in a complete materialization graph is actually the result
of materialization.
Thus, the procedure of materialization can be transformed into the construction of
a complete materialization graph.
In the remainder, we only consider complete materialization graphs and
do not distinguish the term from the
notion of `materialization graphs'.
It should also be noted that there may exist several materialization graphs for a datalog program.

>>>>

\begin{definition}[Parallel Tractability Class]\label{def:ptd}
Given a class $\mathcal{D}$ of
datalog programs sharing the same rule set, we say that $\mathcal{D}$
is a class of datalog programs with \emph{p}arallel
\emph{t}ractability, short a \emph{\texttt{PTD} class},
if there exists an \texttt{NC} algorithm that performs sound and complete materialization for each datalog program
in $\mathcal{D}$. The corresponding class of ontologies of $\mathcal{D}$ is called a
class of ontologies with parallel tractability, short a \texttt{PTO} class.

\textcolor{red}{Suggested revision:} (Parallel Tractability Class). Let $R$ be a set of Datalog
  rules. Materialization w.r.t.\ to $R$ is tractable in parallel, if,
  for every set of facts $\textbf{I}$, there exists an \texttt{NC}
  algorithm that performs sound and complete materialization for
  $P=\langle R, \textbf{I}\rangle$; the set $\mathcal{D}$ of all
  such Datalog programs $P$ is called a \emph{\texttt{PTD} class}.
  The corresponding class of ontologies of $\mathcal{D}$ is called a
  class of ontologies with parallel tractability, short a \texttt{PTO}
  class.
\end{definition}

@@

\begin{definition}[Materialization Graph]\label{def:mg}
A \emph{materialization graph}, with respect to
a datalog program $P=\langle R, \textbf{I}\rangle$, is a directed acyclic graph
denoted by $\mathcal{G}=\langle V, E\rangle$ where,
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item $V$ is the node set and $V\subseteq T_R^{\omega}(\textbf{I})$;
  \item $E$ is the edge set and $E\subseteq T_R^{\omega}(\textbf{I})\times T_R^{\omega}(\textbf{I})$.
\end{enumerate}
For each edge of the form $e(v_1, v_2)$ where $v_1$ and $v_2$ are two nodes,
we say that the node $v_1$ is the parent (node) of $v_2$ and the node $v_2$ is the child (node)
of $v_1$. Further, $\forall H,B_1,\ldots,B_n\in V$, $\mathcal{G}$ satisfies the following conditions:
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item $H$ has a parent node or a child node;
  \item if $H$ has an in-degree of $0$, then $H$ is an original fact of $P$;
  \item $B_1,\ldots,B_n\rightarrow H\in P^*$ is satisfied if $e(B_1, H),\ldots,e(B_n, H)\in E$ and
  $B_1,\ldots,B_n$ are all the parents of $H$.
\end{enumerate}
A materialization graph $\mathcal{G}$ is a \emph{complete materialization graph},
when $\mathcal{G}$ contains all ground atoms in $T_R^{\omega}(\textbf{I})$.

\textcolor{red}{Suggested revision:}
(Materialization Graph).
A \emph{materialization graph}, with respect to
a datalog program $P=\langle R, \textbf{I}\rangle$, is a directed acyclic graph
denoted by $\mathcal{G}=\langle V, E\rangle$ where,
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item $V$ is the node set and $V\subseteq T_R^{\omega}(\textbf{I})$;
  \item $E$ is the edge set and $E\subseteq T_R^{\omega}(\textbf{I})\times T_R^{\omega}(\textbf{I})$.
\end{enumerate}
For each edge of the form $e(v_1, v_2)$ where $v_1$ and $v_2$ are two nodes,
we say that the node $v_1$ is the parent (node) of $v_2$ and the node $v_2$ is the child (node)
of $v_1$.

Further, for each $H,B_1,\ldots,B_n\in V$, $\mathcal{G}$ satisfies the following conditions:
\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
  \item if $e(B_1, H),\ldots,e(B_n, H)\in E$ and
  $B_1,\ldots,B_n$ are all the parents of $H$, then $B_1,\ldots,B_n\rightarrow H\in P^*$;
  \item if $H$ has an in-degree of $0$, then $H$ is an original fact
    of $P$.
\end{enumerate}
A materialization graph $\mathcal{G}$ is a \emph{complete materialization graph},
when $\mathcal{G}$ contains all ground atoms in $T_R^{\omega}(\textbf{I})$.
\end{definition}

@@

The set of nodes in a complete materialization graph is actually the result
of materialization.
Thus, the procedure of materialization can be transformed into the construction of
a complete materialization graph.
In the remainder, we only consider complete materialization graphs and
do not distinguish the term from the
notion of `materialization graphs'.
It should also be noted that there may exist several materialization graphs for a datalog program.
